Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline

### 2017-09-13 draft/response_to_referee.txt

The changes of text related to this comment are: 
    - At the end of Introduction section, we make it clear that we are using AB system, and explain the
      definiton of S/N and Nxsigma detection.  
    - At the end of Section 2.2, we provide more details related to the flux measurements for PSF and CModel
      photometry.  We also refer the reader to Bosch+2017 for more technical details. 
    - Section 6.4: slightly more details about the usePixelWeights option and reference to hscPipe paper.

-----
# 4. CModel algorithm

We agree with the referee that the details of CModel algorithm should be explained better.  However, limited
by the scope of this work, we can not afford to show all the technical details.  As mentioned, we expanded
the description in Section 2.2, and points the readers to the relevant sections in Bosch+2017.

Related to the referee's detailed comments: 
    - We should clearly point out that good performace of cModel photometry is **not** a natural consequence
      of hscPipe, and hscPipe **does not** use GalSim model for galaxy photometry.  These are two completely
      independent software, and SynPipe interacts with them at different stages of the test.  In our tests, the
      synthetic galaxy is described using single-Sersic profile, and is generated using GalSim.  For hscPipe,
      we test CModel photometry which uses very different model assumptions and algorithms for galaxy photometry. 
      It first fits an exponential model to the image with both the shape and the amplitude free. 
      Then it fits a de Vaucouleurs model in the same way.  As last step, it fits both models 
      simultaneously, keeping their ellipses fixed at the results from the previous two fits, and 
      allowing only the two amplitudes to vary.
      Also, in CModel, both of these models are actually described by multi-Gaussian approximations to allow
      fast convolution (e.g. Hog & Lang 2013, Sheldon 2014).  
      The CModel algorithm is not designed to model detailed flux distributions of galaxies accurately, but to
      provide robust and efficient photometry for large amount of galaxies, especially for the poorly-resolved
      and/or low S/N ones.  
    - Although these tests are much more than just "internal consistency", we should mention that 1)
      single-Sersic is still oversimplified model for real galaxies; 2) we assume that there is no error is
      PSF modelling.  These are the limitations of our current tests, and we mention them in the discussion
      section now.  
    - About the central pixel value of Sersic model, please refer to Rowe et al. (2015; 
      http://www.sciencedirect.com/science/article/pii/S221313371500013X) for details of this code.  
      The Section 5.3 and Section 6 explain the integration over pixels and image rendering method in great
      details.  We should point out that synthetic galaxies generated by GalSim have been used for various
      photometric and weak lensing tests.  The algorithms involved should be considered well tested.

We modify the text in the following way: 
    - Section 3.2.1: We point the readers to the relevant sections in Rowe+2015 for more details. 
    - At the end of Section 4.2, we clearly point out that CModel uses very different model assumptions, hence
      we are not only testing the internal consistency of CModel pipeline. 
    - In the discussion section, we add discussion about the limitation of our single-Sersic model test.

-----
# 5. Fig 4 and 9.

We understand the referee's point that random noises from photon statistics could be useful for comparison.

However, as we pointed out earlier, given the design of the camera, observational strategy, and the complex
processes of data reduction, it is impossible to make statistics based on the "input photon number".  
Each pixel on the coadd image consists of information from CCDs with different characteristics and from
separated exposures under different seeing and sky conditions.  These individual exposures are first detrended,
calibrated, background subtracted, and warped before they are coadded using weights that are related to their
variance information.  Given these reasons, it is not realistic to conduct tests using photo number and random
statistical errors as the referee requested.

Meanwhile, we should point out that, since hscPipe (and LSSTpipe in the future) is designed to provide
high-level science-ready results for users, it carefully calibrates the fluxes into AB magnitude system
(absolute flux level), which is physically more meaningful and convinient.  In addition, the hscPipe tries to
provide per-pixel variance information for both single-Visit and coadd images, which includes statistical
uncertainties from sources mentioned and not mentioned by the referee (e.g. correction of brighter-fatter
effect, crosstalk, cosmic-ray removal, number of single-visit images that go into the coadd etc.).
For coadd image, the variance plane provided by hscPipe should be treated as a summary of all statistical
errors, and all the flux error in hscPipe is estimated based on it.  Notice that, for both PSF and CModel
photometry, the estimation of statistical uncertainty is not as straightfoward as for aperture photometry.
And it is not realistic for us to provide independent estimations for each object.

As mentioned earlier, such definition of S/N has been adopted within the HSC SSP collaboration, and has
been used for multiple technical and scientific purposes.  Therefore, we think the S/N plotted in Fig 4 and 
Fig 9 are the most meaningful and useful ones for HSC data user.

We slightly modify the text to explain the definition of S/N and flux error better: 
    - In Section 5.1.1, we explain the PSF flux error in more details.
    - In Section 5.2.1, we explain that for extended objects modeled by CModel, definition of S/N is not as
      straightforward as point sources.  Here we choose to use the CModel flux and flux error to simply
      demonstrate the typical "significancy" of galaxies at fixed magnitude, and show that hscPipe can
      relibaly detect and measure flux for galaxies down to certain magnitude.  We also mention that
      uncertainty of CModel flux is dominated by systematics, and here the flux error only summarises the
      statistical ones.  Therefore, our tests should provide more realistic estimation of average flux error
      (including both systematic and statistical ones) for galaxy at certain magnitude.

-----
# 6. Pure statistical errors in figures.

We agree with the referee that it is very useful to compare with the statistical uncertainties.  However, as
we explained for comment #5,

-----
# 7. Precision and accuracy of PSF magnitude

@@

4. p.13: We think the location and context of this figure makes it very clear that it is for the input
parameters.  Now we state that clearly in the caption.

5. Fig. 4 and 5: has been corrected.

6.

-----
8. P.29

Yes, it should be (i-z) color, and has been corrected.

-----
9. Astrometric Calibration Fig 14

HSC image has pixel scale of 0.168 arcsec/pixel.  We use arcsec as unit since when evaluating the astrometric
calibration for HSC DR1, arcsec or mas are used in Aihara+2017.  Information about pixel scale is added to the
caption of the figure and the text.

### 2017-09-13 draft/ms_v5.tex
Comment 4 5 6 7

        \item \textbf{\ser{}models}:
            \texttt{SynPipe} uses \galsim{} \texttt{v1.4} (\citealt{Rowe2015}) to
            simulate galaxies.
            \galsim{} is an image simulation tool that was designed for the 
            GRavitational lEnsing Accuracy Testing 3 (GREAT3) challenge
            (\citealt{Mandelbaum2014}). 
            Currently, \synpipe{} allows synthetic galaxies to be modeled by a
            single or a double \citet{Sersic1963}
            profile\footnote{More flexible model choices will be added later.}. 
            The \ser{} profile is flexible enough to describe the overall flux
            distributions of galaxies near and far, both early-type or late-type.
            A double-\ser{} model can simulate a galaxy with even more realistic
            structural details (e.g., bulge$+$disk).
            For each \ser{} component, the user needs to provide the magnitude,
            effective radius ($R_{\mathrm{e}}$, in unit of arcsec), \ser{} index
            ($n_{\mathrm{Ser}}$), axis ratio ($\mathrm{b}/\mathrm{a}$), and position
            angle (PA).
            External shear can also be applied to a synthetic galaxy ($g_1$ and
            $g_2$, see \citealt{Rowe2015} for more details).

@@

    For synthetic galaxies, we use single-\ser{} galaxy models of COSMOS galaxies with 
    $I_{F814W} \leq 25.2$ mag. 
    These are models from \citet{Lackner2012} applied to \textit{HST}/ACS images and 
    are included in \galsim{}. 
    Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    Although the single-\ser{} model is not always the best choice for describing 
    galaxies, its flexibility enables it to reasonably describe the flux distribution 
    of most galaxies.
    Also, its simplicity makes it easy for us to diagnose potential problems with the 
    \cmodel{} photometry.
    We exclude a tiny fraction (${\sim}4$\%) of ill-behaved models that have a very 
    high \ser{} index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface 
    brightness ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing objects with problematic photometry (see Appendix \ref{app:qc} for 
    details), we obtain a sample of 58,210 synthetic galaxies with realistic 
    five-band HSC \cmodel{} photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test \hscpipe{}'s general photometric behaviors, 
    but the sample lacks relatively bright galaxies ($i<20.5$ mag).

@@

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses third-order Lanczos interpolation to 
    shift the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainties.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see Bosch\etal (in prep.).

@@

    The \s2n{} for PSF photometry is defined as
    $\mathrm{Flux}_{\mathrm{PSF}}/\mathrm{Flux\_Err}_{\mathrm{PSF}}$ in each band. 
    Figure \ref{fig:star_sn} shows the relationship between stellar magnitude and \s2n{} 
    in the $g$- and $i$-bands. 
    This figure shows the expected \s2n{} of stars as a function of magnitude in the 
    Wide layer.
    The slopes and scatters of these relations are very similar to the ones using 
    real stars on these two \tracts{}.

@@

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$ 
    and $i$ bands, which is consistent with the values found in \citet{HSCDR1} for 
    average seeing conditions.
    However, it is worth reminding HSC data users that the detection
    limit will exhibit spatial variations due to the seeing conditions.
%% ------------------------------------------------------------------------------------ %%

@@

    Fig \ref{fig:cmodel_sn} shows the relation between the input magnitudes and the 
    \s2n{} of synthetic galaxies.  
    The \s2n{} is the ratio  of the \cmodel{} flux and the flux error measured by 
    \hscpipe{}.
    Here the adopted \s2n{} corresponds to the average \s2n{} over entire footprint. 
    The center of the galaxy typically has high \s2n{} than this value.

@@

@@

    Fig \ref{fig:cmodel_cdist} compare the input color--color distributions with the
    recovered ones. 
    The general 2--D distributions are well recovered but the precision is not as 
    good as for stars, especially in the $(i-z)$ vs. $(z-y)$ plane.
    This is expected given the narrow dynamical ranges of these colors in the redder 
    filters.
    The \forced{} \cmodel{} tends to slightly underestimate the $(g-r)$ and $(i-z)$
    colors at the very ``blue'' end, while overestimating the $(i-y)$ color at the very
    red end.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Astrometric Calibration
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_astrometry}
    \end{center}
    \caption{
        Astrometric accuracy for synthetic stars (\textbf{left}) and galaxies
        (\textbf{right}). 
        We show the differences between the input R.A and Dec and the coordinates 
        measured on the \coadd{} images. 
        Filled-contours (filled histogram) and open-contours (solid-line histogram) are 
        for synthetic objects from \tracts{} with good and bad seeing conditions, 
        respectively.
        $\Delta\mathrm{R.A}=0$ and $\Delta\mathrm{Dec}=0$ are marked by dashed lines.     
        It shows that the initial astrometric calibration is already very accurate,  
        and it also suggests that \hscpipe{} can find the centroids of stars and 
        galaxies with precisions that are consistent with the astrometric calibration
        uncertainty under moderately different seeing conditions.
        }
    \label{fig:astrometry}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    Fig \ref{fig:astrometry} shows the distribution of astrometric offsets for 
    synthetic stars and galaxies in \texttt{goodSeeing} and \texttt{badSeeing} 
    \tracts{}.
    We find small systematic offsets at the ${\sim}10$--20 mas level, which is similar 
    to the uncertainties of astrometric calibration quoted in \citet{HSCDR1}.
    The two \tracts{} show different systematic offsets, while the synthetic stars and
    galaxies show coherent offsets in the same \tract{}.
    This shows that the initial astrometric calibration of HSC has very reasonable 
    accuracy.  
    It also suggests that \hscpipe{} can find the centroids of stars and galaxies 
    with precisions that are consistent with the astrometric calibration uncertainty
    under moderately different seeing conditions.

@@

    Fig \ref{fig:blend} shows the $b$ parameter versus the difference between input and
    output values for both magnitudes and colors. 
    For both PSF and \cmodel{} magnitude distributions, we can see clear trends that 
    total fluxes are systematically underestimated for highly blended ($b>0.05$ for 
    stars or $0.1$ for galaxies) objects.
    This effect is particularly strong for stars.

>>>>

        \item \textbf{\ser{}models}:
            \texttt{SynPipe} uses \galsim{} \texttt{v1.4} (\citealt{Rowe2015}) to
            simulate galaxies.
            \galsim{} is an image simulation tool that was designed for the 
            GRavitational lEnsing Accuracy Testing 3 (GREAT3) challenge
            (\citealt{Mandelbaum2014}). 
            Currently, \synpipe{} allows synthetic galaxies to be modeled by a
            single or a double \citet{Sersic1963}
            profile\footnote{More flexible model choices will be added later.}. 
            The \ser{} profile is flexible enough to describe the overall flux
            distributions of galaxies near and far, both early-type or late-type.
            A double-\ser{} model can simulate a galaxy with even more realistic
            structural details (e.g., bulge$+$disk).
            For each \ser{} component, the user needs to provide the magnitude,
            effective radius ($R_{\mathrm{e}}$, in unit of arcsec), \ser{} index
            ($n_{\mathrm{Ser}}$), axis ratio ($\mathrm{b}/\mathrm{a}$), and position
            angle (PA).
            External shear can also be applied to a synthetic galaxy ($g_1$ and
            $g_2$, see \citealt{Rowe2015} for more details).
            \song{ 
            To correctly generate synthetic image of \ser{} or double-\ser{} galaxy
            model, it is not trivial to carefully consider details in pixel 
            integration (especially for central pixel) and image rendering process.  
            Details about these can be found in Section 5 and Section 6 of 
            \citet{Rowe2015}.
            }

@@

    For synthetic galaxies, we use single-\ser{} galaxy models of COSMOS galaxies with 
    $I_{F814W} \leq 25.2$ mag. 
    These are models from \citet{Lackner2012} applied to \textit{HST}/ACS images and 
    are included in \galsim{}. 
    Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    \song{
    It is worth noting that although single-\ser{} profile is still oversimplified model 
    compared to the complexity in real galaxies, they are flexible enough to broadly 
    describe the flux distributions of most galaxies, and test the \cmodel{} algorithm 
    who models galaxy in very different way.
    In addition, single-\ser{} model is simple enough to ensure straightforward 
    diagnostic of potential systematic issues of \cmodel{} algorithm. 
    }
    We exclude a tiny fraction (${\sim}4$\%) of ill-behaved models that have a very 
    high \ser{} index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface 
    brightness ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing objects with problematic photometry (see Appendix \ref{app:qc} for 
    details), we obtain a sample of 58,210 synthetic galaxies with realistic 
    five-band HSC \cmodel{} photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test \hscpipe{}'s general photometric behaviors, 
    but the sample lacks relatively bright galaxies ($i<20.5$ mag).

@@

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses third-order Lanczos interpolation to 
    shift the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainties.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see \citet{Bosch2017}.

@@

    The \s2n{} for PSF photometry is defined as
    $\mathrm{Flux}_{\mathrm{PSF}}/\mathrm{Flux\_Err}_{\mathrm{PSF}}$ in each band. 
    \song{
    The flux error here is estimated based on the per-pixel variance information that
    summaries statistical uncertainties from various sources. 
    However, systematical one like the centroid uncertainty is not taken into account 
    as it should be subdominant even at low-\s2n{}. 
    As mentioned earlier, such definition of PSF \s2n{} is adopted by \citet{HSCDR1} 
    and \citet{Bosch2017}.
    }
    Figure \ref{fig:star_sn} shows the relationship between stellar magnitude and \s2n{} 
    in the $g$- and $i$-bands. 
    This figure shows the expected \s2n{} of stars as a function of magnitude in the 
    Wide layer.
    The slopes and scatters of these relations are very similar to the ones using 
    real stars on these two \tracts{}.

@@

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$ 
    and $i$ bands, which is consistent with the values found in \citet{HSCDR1} for 
    average seeing conditions.
    However, it is worth reminding HSC data users that the detection
    limit will exhibit spatial variations due to the seeing conditions.
    \song{
    In addition, the image warping process introduces covariance across pixels into 
    \texttt{coadd} images, and leads to slightly underestimated the flux error.  
    Therefore, it is useful to compare the statistical results from our tests with the 
    average PSF flux error at fixed input magnitude.  
    }
%% ------------------------------------------------------------------------------------ %%

@@

    Fig \ref{fig:cmodel_sn} shows the relation between the input magnitudes and the 
    \s2n{} of synthetic galaxies.  
    The \s2n{} is the ratio  of the \cmodel{} flux and the flux error measured by 
    \hscpipe{}.
    \song{
    For extended object with all kinds of flux distributions, definition of \s2n{} is 
    not as straightforward as for point sources.  
    Here the adopted \s2n{} corresponds to the average \s2n{} over entire footprint. 
    The center of the galaxy typically has high \s2n{} than this value.    
    We adopt this definition as the main purpose here is to demonstrate the overall 
    ``significancy'' of galaxy down to certain magnitude. 
    }

@@

    \song{
    The \cmodel{} flux error also only considers statistical uncertainties provided by 
    the variance information. 
    For galaxy photometry based on 2-D model, one can expect the error budget is 
    dominated by systematical ones involved in the fitting process.  
    Therefore, statistical tests like the one presented here can provide more insights 
    about the accuracy of \cmodel{} photometry.  
    }

@@

    Fig \ref{fig:cmodel_cdist} compare the input color--color distributions with the
    recovered ones. 
    The general 2--D distributions are well recovered but the precision is not as 
    good as for stars, especially in the $(i-z)$ vs. $(z-y)$ plane.
    This is expected given the narrow dynamical ranges of these colors in the redder 
    filters.
    The \forced{} \cmodel{} tends to slightly underestimate the $(g-r)$ and $(i-z)$
    colors at the very ``blue'' end, while overestimating the \song{
    $(i-z)$ color}
    at the very red end.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Astrometric Calibration
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_astrometry}
    \end{center}
    \caption{
        Astrometric accuracy for synthetic stars (\textbf{left}) and galaxies
        (\textbf{right}). 
        We show the differences between the input R.A and Dec and the coordinates 
        measured on the \coadd{} images. 
        \song{
        HSC \coadd{} image has pixel scale of 0.168\asec{} per pixel. 
        }
        Filled-contours (filled histogram) and open-contours (solid-line histogram) are 
        for synthetic objects from \tracts{} with good and bad seeing conditions, 
        respectively.
        $\Delta\mathrm{R.A}=0$ and $\Delta\mathrm{Dec}=0$ are marked by dashed lines.     
        It shows that the initial astrometric calibration is already very accurate,  
        and it also suggests that \hscpipe{} can find the centroids of stars and 
        galaxies with precisions that are consistent with the astrometric calibration
        uncertainty under moderately different seeing conditions.
        }
    \label{fig:astrometry}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    Fig \ref{fig:astrometry} shows the distribution of astrometric offsets for 
    synthetic stars and galaxies in \texttt{goodSeeing} and \texttt{badSeeing} 
    \tracts{}.
    \song{
    Note that HSC \coadd{} image has pixel size of 0.168\asec{} per pixel. 
    }
    We find small systematic offsets at the ${\sim}10$--20 mas level, which is similar 
    to the uncertainties of astrometric calibration quoted in \citet{HSCDR1}.
    The two \tracts{} show different systematic offsets, while the synthetic stars and
    galaxies show coherent offsets in the same \tract{}.
    This shows that the initial astrometric calibration of HSC has very reasonable 
    accuracy.  
    It also suggests that \hscpipe{} can find the centroids of stars and galaxies 
    with precisions that are consistent with the astrometric calibration uncertainty
    under moderately different seeing conditions.

@@

    Fig \ref{fig:blend} shows the $b$ parameter versus the difference between input and
    output values for both magnitudes and colors. 
    For both PSF and \cmodel{} magnitude distributions, we can see clear trends that 
    total fluxes are systematically underestimated for highly blended \song{($b>0.05$)} 
    objects.
    This effect is particularly strong for stars.
